Radio Emission from Red-Giant Hot Jupiters

### 2015-12-05 reply2referee.txt

  We thank the referee for pointing out that we were not as clear as
  we would like to have been regarding the many differences from and
  advances upon the seminal paper of Ignace (2010) in our original
  submission.  In this resubmission, we have taken the referee's
  comments to heart and have endeavored to trim literature review
  where it is unnecessary, and to highlight the new aspects of our
  work.

## Recommendation:
## 
## My main concern with this paper is that it is much too long for its
## original scientific content. The amendment to the Ignace (2010) result
## is interesting, as is the decay with time of the emission frequency,
## but much of the rest is unnecessary. For example, Appendix B covers
## almost two pages, but is not referred to or used in the paper. I have
## outlined below detailed comments on each section, but my
## recommendation is that the authors reduce the main paper by a factor
## of four by focussing on only the new and relevant results.

  We thank referee for this comment.  We now realize that we had
  previously failed to put our work into proper context.  The main topic
  of our paper is the re-evaluation of radio emission from planets around
  evolved stars. While the possibility of their strong radio emission has
  been addressed in Ignace et al. (2010), there are important updates in 
  this paper, in particular:

  - pointing out the presence of ionizing radition from accretion and
    quantitative evaluation of their ionizing effects. A mechanism to
    ionize otherwise neutral stellar wind has been briefly introduced
    in Ignace et al. (2010), but it is different from the mechanism we
    introduce.
  - taking account of the shrinking of planetary magnetosphere. As a
    result, our estimate of radio emission is order-of-magnitude off
    of their scaling (eq. 20 of their paper).
  - taking account of the plasma frequency of stellar wind, which
    severely affects the detectability of radio signal.

@@

  In the revised manuscript, we tried to make these point clearer in
  Introduction section. The revised parts are marked in bold font.

  Previously, we covered some topics that have already been addressed
  by previous papers, with an intention of making the paper
  self-contained and easy to follow. According to referee's
  suggestion, we attempted to reduce the materials as much as we think
  reasonable.  There are some arguments which we still think are worth
  part of the paper. We describe them in the following, in reponse to
  each of referee's detailed comments.

@@

## Section 3.2
## 
## Given that the authors dismiss the AGB stars because their cutoff
## frequencies are too high (which could have been demonstrated very
## briefly) it seems unnecessary to include these at every step. More
## significantly, the comparison with the "classical" hot Jupiters is not
## valid, either in eqns 43 or 48, as in this case the orbital velocity
## of the planet cannot be neglected when considering the relative
## velocity of the stellar wind particles to the planet (equation 7). The
## stellar rotational velocity is probably small and so can be neglected
## however. The authors should demonstrate that the ram pressure of any
## exoplanetary outflow induced by heating can also be neglected in
## determining the size of the magnetosphere.

  As for the AGB stars, we believe it is important to discuss not only
  potentially detectable signals but also the limitation of our
  ability of detection. In addition, there are some parameter range
  where AGB companions could be detected. Therefore, we think it is
  worth including the discussion of AGB stars.

  As for the comparison with "classical" hot Jupiters, we appreciate
  referee's comments. We estimated the orbital velocity of our
  ``canonical hot Jupiter'' at 0.05 AU to be ~130 km/sec, which can
  contribute 5% increase in the relative velocity. We mentioned in the
  paragraph following equation (43).

  We now added the discussion about the effect of planetary outflow in 
  Appendix A.

@@

  Following the referee's comment, we cut Section 5.1, while some part
  of it are merged with the last section (shown in bold font).  The
  intention of Section 5 was to answer a potential concern about the
  canceling field of far more massive stellar wind than sollar
  wind. We have clarified this motivation and moved it to Appendix.
  Previously we thought Appendix B might be useful for future workers
  (since we often got confused when reading previous papers), but we
  decided to cut it for the sake of brevity, as the referee suggested.

  [DSS: I don't think it should be moved to the Appendix.]

## Section 6:
## 
## The authors note the rareness of evolved stars - it would be useful if
## they included a brief calculation or statement of how many might be
## expected, particularly in the age ranges they consider.

  As far as we understand the referee's feedback, Figure 5 partly
  answers this suggestion, where the expected flux from the known
  system is estimated. We now added explicitly mention the number of
  potential targets above the detection limit. 
  Another way to answer this question we think of is to convolute the 
  planet occurrence rate, noting that not all of these targets are likely
  to harbor a planet. Thus, we also addressed the expected value of
  detection, based on the previously estimated occurrence rate of
  Jupiter-size planets. The method to estimate that number is detailed in
  Appendix D and the main text quotes them.

### 2015-12-06 ms.tex
minor edits

Figure \ref{fig:planetaryB} shows the computed radius ($r_c$) and average density ($\rho_c$) of the dynamo region as function of planetary mass, as well as the heat flux at the outer boundary of the core ($q_o$), the estimated strength of planetary magnetic field ($B$) and the corresponding cyclotron frequency ($\nu_{\rm cyc}$) as functions of planetary mass and the age. 
Substituting equation (\ref{eq:burrowsHeatFlux}) into equation (\ref{eq:Bscaling2}) and given that $r_{\rm dynamo} $ does not change significantly, the magnetic field is approximately:
%%%%%%%%%% 
\begin{equation}
B   \sim   10~{\rm G} \left( \frac{M_p }{M_{p,{\rm J}}} \right)^{1.04} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \label{eq:scalingB}
\end{equation}
%%%%%%%%%%
under this model. 
Reasonably, the resultant values agree with \citet{reiners2010}, who adopted the same scaling law for planetary magnetic field; we show this figure just for completeness. 
Note that the cyclotron frequency of Jovian planets typically falls between 10~MHz and 1~GHz. 
In this regime, there are a number of current and near-future radio wave detectors including GMRT, 
LOFAR, HERA, SKA, and potential upgrades to the VLA (see Section \ref{ss:detectability}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp]
   \plotonesc{cartoon_10Mj_100pc_smooth.pdf}
   \caption{A cartoon of radio emission spectra of a RGHJ with 10$M_{\rm J}$ and the host red giants with 100 $R_{\odot }$.
The spectrum of RGHJ is modeled after Jovian radio spectra, e.g. Figure 8 of \citet{zarka_et_al2004} and Figure 2 of \citet{griesmeier2007a}; contribution from Io-DAM is not shown here.
The spectra of the host red giant are modeled simply by extrapolating observed radio spectra above 1~GHz with power law. }
  \label{fig:cartoon}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

@@

\revise{%
Finally, we note that the radio emission considered in this paper can either be an avenue to learn more about known RGHJs or a tool for discovering new planets, given that traditional exoplanet-discovery methods (specifically, transit and radial-velocity observations) are not very sensitive to planets around RGB or AGB stars.
Once planetary radio emission is detected, it will provide a unique approach for learning about the planetary spin, magnetic field, presence of satellites, and potentially the stellar wind properties.%
}{%
Once planetary radio emission is detected, it in general provides a unique approach for learning the planetary property. As the auroral radio flux is significantly modulated in association with several physical processes described in \S\ref{ss:timevariability}, the time variability is useful to imply the planetary spin rotation period or the presence of satellites. If spectra are obtained, the upper cut-off frequency can tell us the magnetic field strength of the planetary surfaces, which, when combined with constrains on the planetary mass via e.g. radial-velocity observations, tests  the proposed scaling law of planetary dynamo. 
A extra fruit from observations of RGHJs in particular is that the lower cut-off frequency, which presumably corresponds to in situ plasma frequency, could probe of the stellar wind property at the planetary orbit. In addition, the planetary radio power in such an extreme environment as RG's massive stellar wind will be valuable data points to test the empirical ``radiometric Bode's law''. %
We also note that the radio emission of RGHJs may also be used as a tool for discovering/confirming new planets, given that traditional exoplanet-discovery methods (specifically, transit and radial-velocity observations) are not very sensitive to planets around RGB or AGB stars. %
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expected number}
\label{ss:detectability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{cumlative_expected_value.pdf}
   \caption{}
  \label{fig:expected_values}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Figure \ref{fig:planetaryB} shows the computed radius ($r_c$) and average density ($\rho_c$) of the dynamo region as function of planetary mass, as well as the heat flux at the outer boundary of the core ($q_o$), the estimated strength of planetary magnetic field ($B$) and the corresponding cyclotron frequency ($\nu_{\rm cyc}$) as functions of planetary mass and the age. 
Substituting equation (\ref{eq:burrowsHeatFlux}) into equation (\ref{eq:Bscaling2}) and given that $r_{\rm dynamo} $ does not change significantly, the magnetic field is approximately:
%%%%%%%%%% 
\begin{equation}
B   \sim   10~{\rm G} \left( \frac{M_p }{M_{\rm J}} \right)^{1.04} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \label{eq:scalingB}
\end{equation}
%%%%%%%%%%
under this model. 
Reasonably, the resultant values agree with \citet{reiners2010}, who adopted the same scaling law for planetary magnetic field; we show this figure just for completeness. 
Note that the cyclotron frequency of Jovian planets typically falls between 10~MHz and 1~GHz. 
In this regime, there are a number of current and near-future radio wave detectors including GMRT, 
LOFAR, HERA, SKA, and potential upgrades to the VLA (see Section \ref{ss:detectability}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbp]
   \plotoneh{cartoon2_5Mj_100pc_smooth.pdf}
%\begin{figure*}[tbp]
%   \plotonesc{cartoon_10Mj_100pc_smooth.pdf}
   \caption{A cartoon of radio emission spectra of a RGHJ with 10$M_{\rm J}$ and the host red giants with 100 $R_{\odot }$.
The spectrum of RGHJ is modeled after Jovian radio spectra, e.g. Figure 8 of \citet{zarka_et_al2004} and Figure 2 of \citet{griesmeier2007a}; contribution from Io-DAM is not shown here.
The spectra of the host red giant are modeled simply by extrapolating observed radio spectra above 1~GHz with power law. \revise{}{The shadowed region represents the frequency below the plasma frequency of the stellar wind.}}
  \label{fig:cartoon}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\revise{}{A more relistic expected number of detection can be also evaluated by combining these results with the probability for a star to have a planet with certain mass and orbit,  %
Observations of main-sequence stars implies that the number of Jovian planets is about a few to ten percent of the number of stars \citep[e.g.][]{cumming2008,cassan2012}. %
Assuming the same population for RG systems, the expected number of actual detection from RGHJs would be order of unity.  
A more detailed examination using the empirical planet mass function by \citet{cumming2008} is given in Appendix \ref{s:expectedvalue}. 
The analysis indicates that if the detection limit is 10$\mu {\rm Jy}$ the total expected value is about 3. }
%However, as we see in the next subsection, the . and the expected value for massive planets ($M_p>5M_{\rm J}$) which supposedly have relatively high cyclotron frequency is about 1-2.

@@

%\revise{}{The expected values for these targets are }

@@

\revise{%
Finally, we note that the radio emission considered in this paper can either be an avenue to learn more about known RGHJs or a tool for discovering new planets, given that traditional exoplanet-discovery methods (specifically, transit and radial-velocity observations) are not very sensitive to planets around RGB or AGB stars.
Once planetary radio emission is detected, it will provide a unique approach for learning about the planetary spin, magnetic field, presence of satellites, and potentially the stellar wind properties.%
}{%
Once planetary radio emission is detected, it in general provides a unique approach for learning the planetary property. As the auroral radio flux is significantly modulated in association with several physical processes described in \S\ref{ss:timevariability}, the time variability is useful to imply the planetary spin rotation period or the presence of satellites. If spectra are obtained, the upper cut-off frequency can tell us the magnetic field strength of the planetary surfaces, which, when combined with constrains on the planetary mass via e.g. radial-velocity observations, tests  the proposed scaling law of planetary dynamo. 
A extra fruit from observations of RGHJs in particular is that the lower cut-off frequency, which presumably corresponds to in situ plasma frequency, could probe the stellar wind property at the planetary orbit. In addition, the planetary radio power in such an extreme environment as RG's massive stellar wind will be valuable data points to test the empirical ``radiometric Bode's law''. %
We also note that the radio emission of RGHJs may also be used as a tool for discovering/confirming new planets, given that traditional exoplanet-discovery methods (specifically, transit and radial-velocity observations) are not very sensitive to planets around RGB or AGB stars. %
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expected Number of Detection Taking Account of Planet Occurrence Rate}
\label{s:expectedvalue}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{cumulative_expected_value.pdf}
   \caption{The expected number of radio detection of Jovian exoplanets above the given threshold mass. }
  \label{fig:expected_values}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
